---
title: "Income Inequality Analysis"
author: "Ziyue Zhong"
date: "2/28/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r ,echo=FALSE }
library('readxl')
library(reshape2)
#read data
gini <- read_excel('GINI.xls')
edu <-read_excel('Expenditure on education.xls')
agri <- read_excel('Agricultural land %.xls')
GDP_gr <- read_excel('Annual GDP growth %.xls')
co2emis <- read_excel('CO2 per capita.xls')
elec_cons <- read_excel('Electric consu per capita.xls')
exp <- read_excel('Expense(% of GDP).xls')
tech_exp <- read_excel('High-tech Exports.xls')
pop_youn <- read_excel('Population ages 15-64 %.xls')
pop <- read_excel('Population,total.xls')
unemp <- read_excel('Unemployment(% ).xls')

#melt data from 2006-2009
gini <- melt(gini, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
edu <- melt(edu, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
agri <- melt(agri, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
GDP_gr <- melt(GDP_gr, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
co2emis <- melt(co2emis, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
elec_cons <- melt(elec_cons, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
exp <- melt(exp, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
tech_exp <- melt(tech_exp, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
pop_youn <- melt(pop_youn, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
pop <- melt(pop, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))
unemp <- melt(unemp, id.vars = "Country Code", measure.vars = c("2006", "2007","2008","2009"))

#merge all variables into the dataset
dt <- data.frame(gini[c('Country Code','value')], edu['value'],
                   agri['value'],GDP_gr['value'],
                   co2emis['value'],elec_cons['value'],
                   exp['value'], tech_exp['value'],
                   pop_youn['value'],pop['value'],
                   unemp['value'])
#change column name
x <- c('Country code', 'gini','education', 'agriculture', 'GDP_growth', 'co2_emission','elec_con', 'expense', 'tech_exp', 'pop_young', 'pop', 'unemployment')
colnames(dt) <- x


sum(is.null(dt))
#0
#dim(dt)
summary(is.null(dt),)

```

```{r ,echo=FALSE }
#correlation
cor(dt[sapply(dt, is.numeric)],use="pairwise.complete.obs")

#explanatory analysis
par(mfrow=c(1,2))
hist(dt$gini, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
hist(dt$education, col="cadetblue", las=TRUE, cex.lab=1.0, ylab = NULL,cex.lab=1.8,  cex.axis=1.8)
par(mfrow=c(1,2))
hist(dt$agriculture, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
hist(dt$GDP_growth, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
par(mfrow=c(1,2))
hist(dt$co2_emission, col="cadetblue", las=TRUE, cex.lab=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
hist(dt$elec_con, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
par(mfrow=c(1,2))
hist(dt$expense, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
hist(dt$tech_exp, col="cadetblue", las=TRUE, cex.lab=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
par(mfrow=c(1,3))
hist(dt$pop_young, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)
hist(dt$pop, col="cadetblue", las=TRUE,cex.main=1.0, cex.lab=1.8, ylab = NULL, cex.axis=1.8)
hist(dt$unemployment, col="cadetblue",cex.main=1.0, cex.lab=1.8, ylab = NULL,cex.axis=1.8)

```


```{r ,echo=FALSE }
#multiple linear regression

require (leaps)
backward.step <- regsubsets(gini~education+agriculture+GDP_growth+co2_emission+
            elec_con+expense+tech_exp+pop_young+pop+unemployment,
             data = dt,na.action=na.omit, method="backward", nvmax=15)
summary(backward.step)

print('RMSE')
sqrt(summary(backward.step)$rss)
print('R^2 adjusted')
summary(backward.step)$adjr2
print('cp')
summary(backward.step)$cp
print('BIC')
summary(backward.step)$bic

#BIC start to drop at model 9 and cp reach deepest at model 9. RMSE and R2 has little difference between model 9 and model 10. 
#The least important feature is tech_expense.

#Build our first model withour tech_expense
lm.01 <- lm(gini~education+agriculture+GDP_growth+co2_emission+
            elec_con+expense+pop_young+pop+unemployment,
             data = dt,na.action=na.omit)
summary(lm.01)
#Evaluate variable importance using t-statistic
#GDP_growth and expense have p-value > 0.05

#Build second model without GDP_growth and expense
lm.02 <- lm(gini~education+agriculture+co2_emission+
            elec_con+pop_young+pop+unemployment,
             data = dt,na.action=na.omit)
summary(lm.02)

#Try remove the co2_emission
lm.03 <- lm(gini~education+agriculture+
            elec_con+pop_young+pop+unemployment,
             data = dt,na.action=na.omit)
summary(lm.03)
#partial F-test
anova(lm.02,lm.03)

#p value = 0.02396 < 0.05 ----- keep co2_emission

#Selected model: lm.02
#Confidence interval of the model
confint(lm.02, conf.level=0.95)
```



```{r ,echo=FALSE }
#check regression assumption
plot(lm.02)
```

## Including Plots

```{r , echo=FALSE}
#Drop variables don't need
dt[c('GDP_growth','expense','tech_exp')] <- NULL

#Correlation Matrix
cor(dt[sapply(dt, is.numeric)],use="pairwise.complete.obs")

#high correlation between co2_emission and electri_consumption and pop_young ---- remove co2 emission & pop_young

lm.04 <- lm(gini~education+agriculture+
            elec_con+pop+unemployment,
             data = dt,na.action=na.omit)
summary(lm.04)
#check assumption again
plot(lm.04)

#

```

```{r ,echo=FALSE}

pairs(dt[, c('gini', 'education', 'logpop', 'log_elec', 'unemployment')], 
      main="Pairwise Scatterplots of Quantitative Variables", las=TRUE, col="firebrick", pch=19)

boxplot(dt$unemployment)
#Becasue pop is heavily right skewed. I applied log transformantion to the variable. After the transformation, it is normally distributed.
dt$logpop = log(dt$pop)
hist(dt$logpop, col="cadetblue", las=TRUE,cex.main=1.0, ylab = NULL, cex.lab=1.8, cex.axis=1.8)

# According to the scatterplot, elec_con is heavily skewed, I use log to transform it.
dt$log_elec = log(dt$elec_con)

#Create new regression model with transformed elec_con
df <- dt[,colSums(is.na(dt))<nrow(dt)]

lm.05 <- lm(gini~education+agriculture+log_elec+
            logpop+unemployment,
             data = dt,na.action=na.omit)
summary(lm.05)
plot(lm.05)

#Becasue there are constant variance problem, we need to fix outlier problem.
plot(lm.05, which=4, las=TRUE,  cex.axis=1.4, cex.lab=1.4)

influential <- as.numeric(which(!(cooks.distance(lm.05) < 0.05)))
dt_screen <- dt[-influential, ]

cooksd <- cooks.distance(lm.05)
sample_size <- nrow(dt)
influential <- as.numeric(names(cooksd)[(cooksd > (4/sample_size))])
dt1 <- dt[-influential, ]
lm.06 <- lm(gini~education+agriculture+log_elec+
            logpop+unemployment,
            data = dt1, na.action=na.omit, )

plot(lm.06, which=4, las=TRUE,  cex.axis=1.4, cex.lab=1.4)
summary(lm.06)
plot(lm.06)

```

Check regression assumption:
1. x variables are fixed. Data is from worldbank, so we assume no measurement error.
2. constant variance
According to the 'squared std.residual & fitted' value plot, the red line is flat when you go from left to right, so this assumption is satisfied.
3. nonlinearity
According to the 'std. residual & fitted' value plot, there is no obvious trend line, so this assumption is satisfied.
4. normality
The theoretical Quantiles plot is slightly skewed while the violation is not severe.
5. Independence of x
Since each contry can have influence on each other, this assumption may not be true.  
6. No collinearity/multicollinearity
The correlation table shows that there is no collinearity between each variables.

###########################


Final Model: Gini = 1.16999*education-0.10082*algriculture-3.06792*log_elec+39.84998log_pop+0.53149unemployment

